\secrel{6.1 Introducing the Environment  . 26}

The intuition that addresses the first concern is to have the interpreter “look
up” an identifier in some sort of directory. The intuition that addresses the
second concern is to defer the substitution. Fortunately, these converge nicely
in a way that also addresses the third. The directory records the intent to
substitute, without actually rewriting the program source; by recording the
intent, rather than substituting immediately, we can defer substitution; and the
resulting data structure, which is called an environment, avoids the need for
source-to-source rewriting and maps nicely to low-level machine representations.
Each name association in the environment is called a binding.

Observe carefully that what we are changing is the implementation strategy for
the programming language, not the language itself. Therefore, none of our
datatypes for representing programs should change, nor even should the answers
that the interpreter provides. As a result, we should think of the previous
interpreter as a “reference implementation” that the one we’re about to write
should match. Indeed, we should create a generator that creates lots of tests,
runs them through both interpreters, and makes sure their answers are the same.
Ideally, we should prove that the two interpreters behave the same, which is a
good topic for advanced study.

Let’s first define our environment data structure. An environment is a list of
pairs of names associated with...what?
\note{One subtlety is in defining precisely what “the same” means, especially
with regards to failure.}

\DoNow{
A natural question to ask here might be what the environment maps names to. But
a better, more fundamental, question is: How to determine the answer to the
“natural” question?
}

Remember that our environment was created to defer substitutions. Therefore, the
answer lies in substitution. We discussed earlier (Oh Wait, There’s More!) that
we want substitution to map names to answers, corresponding to an eager function
application strategy. Therefore, the environment should map names to answers.
\lsts{src/6/6_1_1.rkt}{rkt}