\secrel{16.1 Contracts as Predicates   . 181}

It is therefore easy to see how to implement simple contracts. A contract
embodies a predicate. It consumes a value and applies the predicate to the
value. If the value passes the predicate, the contract returns the value
unmolested; if the value fails, the contract reports an error. Its only
behaviors are to return the supplied value or to error: it should not change the
value in any way. In short, on values that pass the predicate, the contact
itself acts as the identity function.
\note{In what follows we will use the language \#lang plai, for two reasons.
First, this better simulates programming in an untyped language.
Second, for simplicity we will write type-like assertions as contracts, but in
the typed language these will be flagged by the type-checker itself, not letting
us see the run-time behavior. In effect, it is easier to “turn off” the type
checker. However, contracts make perfect sense even in a typed world, because
they enhance the set of invariants that a programmer can express.}

We can encode this essence in the following function:
\lsts{16/1/1.rkt}{rkt}
Here’s an example contract:
\lsts{16/1/2.rkt}{rkt}
(In a typed language, the number? check would of course be unnecessary because
it can be encoded—and statically checked!—in the type of the function using the
contract.) Suppose we want to make sure we don’t get imaginary numbers when
computing square roots; we might write
\lsts{16/1/3.rkt}{rkt}
In many languages assertions are written as statements rather than as
expressions, so an alternate way to write this would be:
\lsts{16/1/4.rkt}{rkt}
(In some cases this form is clearer because it states crisply at the beginning
of the function what is expected of the parameters. It also enables parameters
to be checked just once. Indeed, in some languages the contract can be written
in the function header itself, thereby improving the information given in the
interface.) Now if real-sqrt- 1 or real-sqrt-2 are applied to 4 they produce 2,
but if applied to -1 they raise a contract violation error.
