\clearpage\label{krishna}
\begin{verbatim}
From: Shriram Krishnamurthi <sk@cs.brown.edu>
Date: Sat, 7 Jan 2017 17:30:53 -0500
To: Dmitry Ponyatov <dponyatov@gmail.com>
\end{verbatim}
% Subject: Re: What legal limitations does your PLAI book have on translation ?

My apologies for not replying sooner. Your message got lost in the mess of my
inbox as my fall semester was in full force and I was traveling, and I only just
saw it. Please accept my apologies.

\bigskip
Thank you for your interest in translating PLAI\,!

\bigskip
I own the copyright to that material, so I have the right to allow a
translation. I am happy to let you translate it.

\bigskip
It is important to me that readers not assume that everything in the translated
version is as accurate as in the original. Unfortunately I'm unable to read
Russian, so I can't even read it and confirm that it is correct. Therefore, I
ask that you

\begin{enumerate}[nosep]
  \item 
Provide a link to the original document
\verb|[|which you seem to have done\verb|]|
  \item 
Say at the top of each page that the translation has not been checked by the
original author
  \item 
State that the correct, definitive version is at the original link
\end{enumerate}

Please confirm that you can comply with this. On these conditions, you are
welcome to continue and list yourself as the translator (but please also
attribute me as the original author).

\bigskip
Note that I would prefer you consider the second edition of the book, which is
newer and has much better prose and improved organization.

\bigskip
Thanks again. Please keep me posted on your progress.

\bigskip\copyright\ Shriram
